Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-20 arxiv/diff_cov.tex
wip

\section{The Different Covariates Setting}
As mentioned in Section \ref{sec_prelim}, the optimization objective \eqref{eq_mtl} is non-convex in general.
Therefore, we consider an arbitrary local minimum $B, W_1, W_2$, in particular when $B = \hat{B}(W_1, W_2)$ satisfies the local optimality condition.
We derive deterministic conditions to compare the prediction loss of the local minimum to single-task learning.
Our key insight is a bias-variance decomposition of the expected prediction loss of $\hat{\beta}_2^{\MTL} = \hat{B} W_2$.
Using the local optimality condition of $B$, we obtain that
\begin{align*}
	 \hat{B}(W_1, W_2) &= (W_1^2 X_1^{\top}X_1 + W_2^2 X_2^{\top}X_2)^{-1} (W_1 X_1^{\top}Y_1 + W_2 X_2^{\top}Y_2)\\
	&= \frac{1}{W_2} \left( \frac{W_1^2}{W_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(\frac{W_1}{W_2} X_1^{\top}Y_1 + X_2^{\top}Y_2\right) \\
	&= \frac{1}{W_2}\left[\beta_2 + \left(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}\bigbrace{X_1^{\top}X_1\left(\frac{W_1}{W_2}\beta_1 - \frac{W_1^2}{W_2^2} \beta_2\right) + \left(\frac{W_1}{W_2} X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2\right)}\right].
\end{align*}
Recall that $\hat{\beta}_i^{\MTL} = \hat{B} W_i$. Therefore, we have
\begin{align}
	\exarg{\epsilon_1, \epsilon_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}
	=&~ \frac{W_1^2}{W_2^2} \bignorm{\Sigma_2^{1/2}(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)}^2 \label{eq_bias_2task} \\
			&+~  \sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }. \label{eq_var_2task}
\end{align}
Equation \eqref{eq_bias_2task} is the bias of $\hat{\beta}_t^{\MTL}$ and
equation \eqref{eq_var_2task} is the variance of $\hat{\beta}_t^{\MTL}$.
%minus the variance of $\hat{\beta}_t^{\STL}$, which is always negative.
Comparing the above to single-task learning, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

>>>>

\section{The Different Covariates Setting}
As mentioned in Section \ref{sec_prelim}, the optimization objective \eqref{eq_mtl} is non-convex in general.
Therefore, we consider an arbitrary local minimum $B, W_1, W_2$, in particular when $B = \hat{B}(W_1, W_2)$ satisfies the local optimality condition.
We derive deterministic conditions to compare the prediction loss of the local minimum to single-task learning.
